---
title: "P8105 Homework 1"
author: "Claire Liu"
date: "2026-09-26"
output: github_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Problem 1

Load the penguins dataset from palmerpenguins.

```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables.
The variable names are `r paste(names(penguins), collapse = ", ")`.
The species included are `r paste(unique(pull(penguins, species)))`.
The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` mm.

Next, create a scatterplot of flipper length against bill length, with points colored by species.

```{r warning=FALSE}
penguin_plot = 
  ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

penguin_plot
```

Then, save the scatterplot to the project directory.

```{r warning=FALSE}
ggsave('penguin_scatterplot.pdf', penguin_plot, height = 4, width = 6)
```



