---
title: "P8105 Homework 1"
author: "Claire Liu"
date: "2026-09-26"
output: github_document
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Problem 1

Load the penguins dataset from palmerpenguins.

```{r}
data("penguins", package = "palmerpenguins")
```

The `penguins` dataset contains `r nrow(penguins)` observations and `r ncol(penguins)` variables.
The variable names are `r paste(names(penguins), collapse = ", ")`.
The species included are `r paste(unique(pull(penguins, species)))`.
The mean flipper length is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE), digits = 2)` mm.

Next, create a scatterplot of flipper length against bill length, with points colored by species.

```{r warning=FALSE}
penguin_plot = 
  ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

penguin_plot
```

Then, save the scatterplot to the project directory.

```{r warning=FALSE}
ggsave('penguin_scatterplot.pdf', penguin_plot, height = 4, width = 6)
```



## Problem 2

Create dataframe contain numeric, logical, character, and factor variables.

```{r}
problem2_df = 
  tibble(normal_sample = rnorm(10),
         sample_positive = normal_sample > 0,
         character_vector = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
         factor_vector = factor(c("A", "A", "A", "A", "B", "B", "B", "C", "C", "C")))

problem2_df
```

Next, try to take the mean of each variable in the dataframe.

```{r}
mean(pull(problem2_df, normal_sample))
mean(pull(problem2_df, sample_positive))
mean(pull(problem2_df, character_vector))
mean(pull(problem2_df, factor_vector))
```

The mean works for the numeric and logical variables. For the logical variable, R treats `TRUE` as 1 and `FALSE` as 0, so the mean represents the proportion of values that are `TRUE`. 
The mean does not work for the character or factor variables.

Then, convert the logical, charactor, and factor variables to numeric variables using `as.numeric`.

```{r results='hide', warning=FALSE}
as.numeric(pull(problem2_df, sample_positive))
as.numeric(pull(problem2_df, character_vector))
as.numeric(pull(problem2_df, factor_vector))
```

Converting the logical variable to numeric changes `TRUE` to 1 and `FALSE` to 0. This helps explain why a mean can be calculated for the logical variable. 

Converting the character variable to numeric produces `NA` values, which means character values cannot be converted to numbers. 

Converting the factor variable to numeric produces integer values corresponding to its factor levels.



